Rule learning for verification from temporal data

### 2016-09-04 Travis/secs/intro.tex
more

In our previous work we built a prototype to verify configuration files from a set of correct examples~\cite{Santolucito2016}.
The tool, ConfigC, takes as a training set, a set correct files and builds a language model.
This model is then used to check if new a configuration file adhears to those rules.
If a file is incorrect, the tool will identify the location of error, and report what is incorrect.
For example, ConfigC might output \texttt{Ordering Error: Expected "extension mysql.so" before "extension recode.so"}.
While this prototype works well, it can only learn from correct examples and has a high false positive rate (marking correct files as incorrect).

In order to extend the prototype, in this paper we provide a new description of ConfigC in terms of version space learning.
This description will allow us understand how the justifications are produced in ConfigC.
We can the extend ConfigC with a new algorithm to decrease the false positive rate, while maintaning the ability to provide justification.
%and show that the results are garunteed to be complete.

%Although the results are complete, the approach used previously produces a high false positive rate.
We propose an extension to ConfigC that allows it to handle both correct and incorrect files in the training set.
Additionally, we extend ConfigC with the ability to use sstructue within the training set.
While the training set from ConfigC was a random sampling of configuration files, many other training sets have a structure based on version history.
%The algorithm we propose takes advantage of this commonly availible, but underutilized structure of training sets for software analysis.
%Because most code does not exist in isolation, but changes over time with development, any training set of code from a verison control system, like Github, has a rich temporal structure. 
This structure is simply a partial order, and can be used by our proposed algorithm for more effective learning.
%A learning algorithm that uses the temporal structure of code to create sound classification with a low false positive rate can be used in at least the above listed software analysis techniques.

@@

A recent usage study of TravisCI found that 15-20\% of failed TravisCI builds are due to "errors" - which means the configuration file was malformed and the software could not even be built \cite{API}.
Using the data from \cite{API}, we can also learn that since the start of 2014, approximately 88,000 hours of server time was used on TravisCI projects that resulted in an error status.
This number not only represents lost server time, but also lost developer time, as programmers must wait to verify that their work does not break the build.
If these malformed projects could be quickly statically checked on the client side, both TravisCI and its users could benefit.

>>>>

In our previous work we built a prototype to verify configuration files from a set of correct examples~\cite{Santolucito2016}.
The tool, ConfigC, takes as a training set a set of correct configuration files and builds a set of rules describing a language model.
This model is then used to check if new a configuration file adhears to those rules.
If a file is incorrect, the tool will identify the location of error, and report what is incorrect.
For example, ConfigC might output \texttt{Ordering Error: Expected "extension mysql.so" before "extension recode.so"}.
While this prototype's results are promising, the main weakness are that the learning can only use correct configuration files.
Additionally, there is a relatively high false positive rate (marking correct files as incorrect).

In order to extend the prototype, in this paper we provide a new description of ConfigC in terms of version space learning.
This description will allow us to understand how the justifications are produced in ConfigC.
We can then extend ConfigC with a new algorithm to decrease the false positive rate, while maintaning the ability to provide justification.
%and show that the results are garunteed to be complete.

%Although the results are complete, the approach used previously produces a high false positive rate.
We propose an extension to ConfigC that allows it to handle both correct and incorrect files in the training set.
Additionally, we extend ConfigC with the ability to use structue within the training set.
While the training set from ConfigC was a random sampling of configuration files, many other training sets have a structure based on version history.
%The algorithm we propose takes advantage of this commonly availible, but underutilized structure of training sets for software analysis.
%Because most code does not exist in isolation, but changes over time with development, any training set of code from a verison control system, like Github, has a rich temporal structure. 
This structure is simply a partial order, and can be used by our proposed algorithm for more effective learning.
%A learning algorithm that uses the temporal structure of code to create sound classification with a low false positive rate can be used in at least the above listed software analysis techniques.

@@

A recent usage study~\cite{API} of TravisCI found that 15-20\% of failed TravisCI builds are due to "errors" - which is the TravisCI code used to mean the configuration file was malformed and the software could not even be built.
Using the data from \cite{API}, we can also learn that since the start of 2014, approximately 88,000 hours of server time was used on TravisCI projects that resulted in an error status.
This number not only represents lost server time, but also lost developer time, as programmers must wait to verify that their work does not break the build.
If these malformed projects could be quickly statically checked on the client side, both TravisCI and its users could benefit.

Our main contributions are then as follow:

\begin{enumerate}

\item We give a description of our prototype tool, ConfigC, in the context of version space learning.
\item We propose a new algorithm for ConfigC to handle both incorrect and correct training data, as well as internal structure of a training set.
\item We propose a real-world application of this approach, and outline the steps needed for an implementation.

\end{enumerate}

### 2016-09-04 Travis/secs/temporal.tex
more

ConfigC produces clean error messages, but has a high false positive rate.
However, ConfigC only analyzes correct configuration files, and does not consider any ordering between these correct files.
We present an extension to ConfigC to decrease the false positive rate in ConfigC by learning on both correct and incorrect examples, as well as temporal structure of these examples.
Our solution is to buid a logical formula representing the entire history of examples, and use a SMT solver to find a classification model.

@@

Since we now consider both positive and negative examples, we will denote the build status of $P_t$ with $S(P_t)$.
In this application, we consider all non-erroring build status to be passing, denoted $Pass$ and otherwise $Err$.
For brevity, we denote sequences of build statuses with the following notation:
\begin{align*}
  S(P_t)=Pass \land S(P_{t+1})=Err \implies S(P_{t,t+1}) = PE
\end{align*}

In contrast with ConfigC, we now consider both positive and negative examples and so must introduce the \textit{general boundary}.
The general boundary is the dual of the specific boundary, and is the most relaxed requirement for a positive classification.
We denoted specific boundary as the set of necessary relations $Nec$, and now denote the general boundary as the set of breaking relations $Br$.
With this notation, we can formally express the requirement that the program summary is sufficently detailed.
\begin{align}
  \forall S(P_t)=Err, \exists r \in M(P_t), r \in Br \label{eq:E1}
\end{align}

>>>>

ConfigC produces clean error messages, but has a high false positive rate.
However, ConfigC only analyzes correct configuration files, and does not consider any ordering between these correct files.
We present an extension to ConfigC to decrease the false positive rate in ConfigC by learning on both correct and incorrect examples, as well as temporal structure of these examples.
This extension will buid a logical formula representing the entire history of examples, and use a SMT solver to find a classification model.

@@

Since we now consider both incorrect and correct data, we will denote the build status of $P_t$ with $S(P_t)$.
In this application, we consider all non-erroring build status to be passing, denoted $Pass$ and otherwise $Err$.
For brevity, we denote sequences of build statuses with the following notation:
\begin{align*}
  S(P_t)=Pass \land S(P_{t+1})=Err \implies S(P_{t,t+1}) = PE
\end{align*}

In contrast with ConfigC, we now consider both incorrect and correct and so must introduce the \textit{general boundary}.
The general boundary is the dual of the specific boundary, and is the most relaxed requirement for a positive classification.
We denoted specific boundary as the set of necessary relations $Nec$, and now denote the general boundary as the set of breaking relations $Br$.
With this notation, we can formally express the requirement that the program summary is sufficently detailed.
\begin{align}
  \forall S(P_t)=Err, \exists r \in M(P_t), r \in Br \label{eq:E1}
\end{align}

### 2016-09-04 Travis/secs/vspace_verify.tex
newintro
more

ConfigC~\cite{Santolucito2016} learns a classification model from a set of correct configuration files, then reports misconfigurations.
Specifically, ConfigC uses version space learing to identify errors in configuration files for MySQL.
Version space learning builds a logical constraint model for binary classification, or membership in a set~\cite{mitchell82}.
Traditional version space learning builds a model using a series of disjunctions from a set of predefined hypotheses.
By instead restricting the model to a series of conjunctions, ConfigC can not only flag misconfigurations, but also give the points of failure for non-membership.

@@

It is assumed if a file is correct, all relations in that file are in the set of necessary relations for any other file to be correct ($Correct(C) \implies \forall r \in M(C), r \in Nec$).
In this way, the initial model is built by creating the strongest conditions for a correct file, called the \textit{specific boundary} in version space learning.
This model is then iteratively relaxed as more examples are seen, a process called \textit{candidate elimination}.
Since we maintain the strongest condition for correctness, this approach will identify many correct files as incorrect, called a high false positive rate.
This is a problem when a user is then asked to manual review many files for errors, when the errors are not truly a problem.

When the system is run on a new file, one of the relations from the templates will not be satisfied.
To provide an error message,the system only needs to report which relation is not satisfied.
For example, ConfigC might output \texttt{Ordering Error: Expected "extension mysql.so" before "extension recode.so"}.

>>>>

Version space learning builds a logical constraint model for binary classification, or membership in a set~\cite{mitchell82}.
Traditional version space learning builds a model using a series of disjunctions from a set of predefined hypotheses.
In ConfigC, we instead restrict the model to a series of conjunctions.
This way, ConfigC can not only flag misconfigurations, but also give the points of failure for non-membership.
We now show in detail how this process works.

@@

It is assumed if a file is correct, all relations in that file are in the set of necessary relations for any other file to be correct ($Correct(C) \implies \forall r \in M(C), r \in Nec$).
In this way, the initial model is built by creating the strongest conditions for a correct file, called the \textit{specific boundary} in version space learning.
This model is then iteratively relaxed as more examples are seen, a process called \textit{candidate elimination}.
Since we maintain the strongest condition for correctness, this approach will identify many correct files as incorrect, which we call a high false positive rate.
This is a problem when a user is then asked to manually review many files for errors, when the files are in fact correct.

When the system is run on a new file, one of the relations from the templates will not be satisfied.
To provide an error message, the system only needs to report which relation is not satisfied.
